---
title: "R Notebook"
output:
  html_document:
    highlight: haddock
    theme: paper
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse);library(tidygraph);library(tidytext);library(ggraph);library(visNetwork)
library(paletteer)
knitr::opts_chunk$set(
  fig.width=6,fig.asp = 0.618,dpi=300,echo=FALSE,message=FALSE,warning=FALSE
)
motifs <- feather::read_feather("data/motifs.feather")
source("prep/motif_network.R")
```

```{r} 
ntwk %>% activate(nodes) %>% as_tibble() %>% View()
```

# Trace a Concept in the Current Hierarchy

## Where does a single term/phrase show up across the entire 'tree' of motifs?

Given a particular term or phrase, can we show all nodes which match, as well as the parent nodes of these, branching back to the root?

How might this work with more complex constellations of terms, like `rise(s) from dead|resurrect|back to life`, or `goddess|serpent` related to Marija Gimbutas' work, or `transfigur|transform|becomes|turns into`?

```{r}

```


Can we display related search output of similar terms/concepts for comparison?  For instance, `resurrection` and `reincarnation`?

```{r}

```


Options for display/exploration:


- Subset the graph dataframe and re-draw a new dendrogram with only the filtered nodes

In order to identify the portion of the overall tree which includes a set of terms, we need to identify not only the nodes which contain those search terms but also their parent nodes, all the way back to the root.^[This is done via the `trace_to_root()` convenience function defined here.]  This allows us to keep the main trunk of the tree intact and connect the disparate items that a search turns up.

```{r}
source('funs/trace_to_root.R')
```

```{r subset_trace}
viz_df <-
  ntwk %>% activate(nodes) %>% 
  filter(row %in% trace_to_root("blood"))

viz_df %>% activate(nodes) %>% as_tibble %>% View()
```


```{r}
viz_df %>%
  ungroup() %>%
  ggraph(layout = 'dendrogram', circular = T) + 
  geom_edge_diagonal(alpha = 0.25) +
  geom_node_point(
    aes(
      colour = name,
      filter = leaf,
      size= level * 0.5, 
      alpha=0.2
    )
  ) +
  geom_node_text(
    aes(
      x = x*1.15, y=y*1.15, 
      filter = leaf, 
      label = str_trunc(as.character(motif_name),20,"right"), 
      angle = -((-node_angle(x, y)+90)%%180)+90, 
      hjust ='outward', 
      colour = name
    ), 
    size=2.5, alpha=1
  ) +
  # scale_fill_paletteer_d(viridis, inferno) +
  coord_fixed() +
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0),"cm")
  ) +
  expand_limits(x = c(-2, 2), y = c(-2, 2))

```

# Compare Traditions

## Identify traditions

*How to extract structured fields from* `notes` *field for each country tagged?*

Identify all traditions referenced in `note` field for tagging
Make logical variables detecting presence of tradition in notes field (potentially using `purrr::map_`)

```{r}
traditions <-
  df %>%
  unnest_tokens(word, notes) %>%
  group_by(word) %>% summarize(n = n()) %>%
  filter(
    word %in% c(
      "india","hindu","indian","american",
      "irish","celtic","ireland",
      "arab","assyrian","turkish",
      "jewish","hebrew","palestine","icelandic","icel","iceland","norse",
      "italian","sicilian","africa","african","zulu","angola","rhodesia","cameroon","congo","nigeria","egypt","egyptian","yoruba","swahili","madagascar","togo","masai","surinam","bantu","liberia","senegambia",
      "chinese","spanish","panama","columbia","ecuador",
      "u.s","missouri","california","calif","georgia","zuñi","ojibwa","cherokee","blackfoot","apache","pueblo","cheyenne","iroquois","mexican","alaska","dakota","arapaho","choctaw","penobscot","hiawatha","tarahumare",
      "korean","indies","nassau","zanzibar",
      "indonesia","indonesian","fiji","maori","samoa","oceanic","oceania","melanesia","melanesian","polynesian","polynesia","micronesia","micronesian","honolulu","hawaiian","sumatra","malay","philippines","java","papuan","raratonga",
      "canadian","canada","estonian","esthonian","swedish","scotch","scottish","britain","british","prussian",
      "england","greek","lithuanian","latvian","rumanian","romania","slavic","eskimo","kodiak","austrian",
      "french","france","danish","russian","bulgarian","slovenian","tahiti","tahltan","rarotonga","jamaica","flemish","burmese","circassian",
      "siberian","swiss","scotland","persian","negro","brazil","peru","argentina","aztec","inca","amazon","hebrides","creole","haitian",
      "finnish","norwegian","hawaii","japanese","buddhist","german",
      "greenland","english","breton","hungarian","korea",
      "wales","welsh","philippine","babylonian","armenian","australian","australia","araucanian","antigua","cyprus",
      ###
      "ikeda","malalasekera","fox","eddic","fang","guinea",
      "nasreddin","anesaki","easter","cook","verde","bering","labrador","wakweli","jijena","crow","snorra","marshall","zealand","nederlandsche","nederlandsch","tlingit","ashanti","teton","yagua","sherente","tehuelche","tenetehara","berbères"
    )
  )

```


```{r fig.height=8}
traditions %>%
  filter(n >= 100) %>%
  ggplot(aes(x=n, y=fct_reorder(word,n,.desc=T))) +
  geom_segment(
    aes(
      x = 0, y = fct_reorder(word,n,.desc=T),
      xend = n, yend = fct_reorder(word,n,.desc=T)
    ),
    color = "gray50"
  ) +
  geom_point() +
  labs(
    x = "Number of tags", y = "Tradition", 
    title = "Motifs from various traditions",
    subtitle = "showing number of instances mentioned (where n > 100)",
    caption = "from the Thompson Motif Index"
  ) +
  theme_minimal()
```


## Which motifs are common across multiple traditions?

Visualize relationships across traditions using hierarchical edge bundling diagram (i.e. `geom_conn_bundle()`)

```{r}

```

Visualize country coverage using coordinates of central city from a given civilization (*and radius based on extent of historical scope?*)

## Given a specific tradition, which others share the greatest number of motifs?

# Uncover Themes

## Clustering motifs to find archetypes

How might we start at the most granular level and develop new categories other than Thompson's?

Some options:

- Topic models (LDA) based on `description` field (and perhaps `notes`)
- String similarity metrics using `stringdist` package

Issues to address:

- How to create/identify clusters?
- How to name them automatically/intuitively

## Needed cleaning steps

In order to identify similarities, it may help to:

- Replace/expand to common synonyms
- Use stem words (lemmatization)

